### 2016-08-15 rebuttal.tex

\begin{verbatim}
>>> trigsimp (exp( Matrix(2, 2, [0, -y, y, 0])) )
\end{verbatim}
fails to recognize cos and
sin. Am I expecting too much?
\begin{solution}
  This indeed does not work. The implementation in \verb|trigsimp| primarily
  thus far has focused on transforming trigonometric functions into other
  trigonometric functions. The ability to simplify complex exponentials is
  something that we would like to work.
  \href{https://github.com/sympy/sympy/issues/11459}{Issue 11459} in our
  public issue tracker tracks this feature.
\end{solution}
\question Section 3.6: are eigenvalues and singular
values included? If not, why not?
\begin{solution}
  % TODO
\end{solution}
\question Like one of the referees, I expected to
see MPFR mentioned in Section 4.1, at least to mention the pros and cons and
say why it isn’t used.
\begin{solution}
  % TODO
\end{solution}
\question Section 4.1: please state whether the syntax for
functions in mpmath is identical, or not, to that in Sympy.
\begin{solution}
  % TODO
\end{solution}
\question Unless I have
missed something, a weakness of the paper is that speed is mentioned only
twice, on lines 496 and 666. My assumption is that SymPy is slow compared with
its commercial competitors. Please comment on the speed issue, and not just in
the conclusions.
\begin{solution}
  % TODO
\end{solution}
\question Editors are needed for [20].
\begin{solution}
  % TODO: I do not know how to make bibtex (with apalike) show the editor
  % here.
\end{solution}
\end{questions}

\section{Reviewer 1 (James Davenport)}
\subsection{Basic reporting}
\begin{questions}
\question Generally good. A part that confused me is the assertion (footnote 3) that "If
A and B are Symbols created with commutative=False then SymPy will keep A·B
and B·A distinct." Does that mean that BOTH of them must be created this way,
and that A and x (if x is created normally) will commute? Is there any way to
declare commutators? How does one guarantee that other pieces of code, e.g.
Gaussian elimination, respect non-commutativity?
\begin{solution}
  We have clarified the footnote. Both expressions must be set as
  \texttt{commutative=False}. Internally, in \texttt{Mul}, the ``commutative
  part'' of an expression is pulled out to the front and canonically ordered,
  and the ``noncommutative part'' is not reordered.

Commutators are implemented in the \texttt{sympy.physics.quantum} module and
are discussed in section 5.2.

Different algorithms do generally require special consideration for
noncommutative expressions to be correct. If an algorithm implicitly assumes
that expressions commute, it may return incorrect results when given a
noncommutative expression.
\end{solution}

### 2016-08-15 rebuttal.tex

The measure parameter of the simplify function lets the user specify the Python function used to determine how

\begin{solution}
We have changed the sentence as suggested.
\end{solution}
\question Section 4.1
Should mpmath be cited here?

\begin{solution}
We have added a citation for mpmath.
\end{solution}
\end{questions}

\section{Reviewer 3 (Vedran Sego)}

\begin{questions}

\question Explaining what Python is (lines 70–72) should go before talking about SymPy as a “CAS written in Python”. Further, the paper assumes a moderate familiarity with Python (for example, Python’s console, OOP, and exceptions), and this should be specified. There should be a short note on the used Python console (>>> is the prompt, with the results of computation following immediately in the lines after it). The citation [25] from line 65 should be moved next to “Python” in line 70.
\begin{solution}

\end{solution}

\question Line 73 has outdated information. Sage was renamed to SageMath and it no longer aims only at pure mathematics but also at algebra, numerical analysis, etc. The reference [40] should be replaced by a more up-to-date one.
\begin{solution}

\end{solution}

\question The plural “CASs” is usually written as “CASes” or “CAS’s”, with the latter being somewhat problematic due to it looking like it implies posses- sion.
\begin{solution}

\end{solution}

\question Line 88 mentions “printers”, but it doesn’t state what they are, which is confusing for those readers that are yet to learn the concept in section 3.4.
\begin{solution}

\end{solution}

\question Also in line 88, Jupyter’s citation [30] is actually about IPython and should be replaced by a more up-to-date version.
\begin{solution}

\end{solution}

\question The word “software” in line 91 is ambiguous; “library” or “package” would make a better choice.
\begin{solution}

\end{solution}

\question Lines 91–96: “we discuss/look at/etc” is the preferred form, instead of “section discusses/looks at/etc”.
\begin{solution}

\end{solution}

\question The paragraph in lines 103–105 should be moved to the introduction, and the footnote from line 104 should be added to that paragraph as a full-blown sentence, expanded by all the relevant technical information (Python version, OS,...). Given that the end of life for Python 2 is 2020., a comment on whether all the presented examples work in both Python 2 and 3 should be included as well. Further, emphasise that wildcard imports, “import *”, should almost never be used in programs (see PEP 8, the item “wildcard imports”). The same goes for the import mentioned in lines 484–487.
\begin{solution}

\end{solution}

\question Line 119 should be removed, as it is basically a copy of the previous line.
\begin{solution}

\end{solution}

\question In line 121 the word “stored” should be replaced by “used as keys”.
\begin{solution}

\end{solution}

\question Whatdotheauthorsmeanby“therebypermittingfeaturessuchascaching” in line 122? Caching can be done for mutable types as well, just not through hashing. 2
\begin{solution}

\end{solution}

\question There is no need to repeat “(CAS)” in line 124, as it was already given in line 64.
\begin{solution}

\end{solution}

\question In the same line, the word “represents” should be replaced by “stores”.
\begin{solution}

\end{solution}

\question In line 184, “symbols are” should be replaced by “t is” (the general rule is already given in line 179).
\begin{solution}

\end{solution}

\question The code in line 210 should be made into its own line (like a displaymath formula), for typesetting reasons and better readability.
\begin{solution}

\end{solution}

\question The footnote 4 should be moved from line 221 to line 222, right after “Basic”.
\begin{solution}

\end{solution}

\question The part “which defines some basic methods for symbolic expression trees” should be removed from line 222, as it was already given in line 130.
\begin{solution}

\end{solution}

\question In line 225, the sentence “Not all SymPy classes are subclasses of Expr.” sounds confusing as a reader new to SymPy wouldn’t expect, for example, symbols to inherit “Expr”. It would be better to expand this, for example “Most of the SymPy classes (including Symbol) are subclasses of Expr, but there are exceptions to this rule”.
\begin{solution}

\end{solution}

\question The title “Features” in line 276 is ambiguous, as a “feature” has no precise meaning in Python (or even software libraries in general). It should be replaced by “Packages and modules” or a similar more precise wording. The same goes for ‘feature” in most other places in the paper (for exam- ple, the caption of Table 1. and line 495). It would be very useful to also include actual names of the packages/modules in Table 1., as well as in any section covering those packages/modules.
\begin{solution}

\end{solution}

\question The sets support listing is unnaturally split in two by the “This includes...” sentence which would fit better in parentheses.
\begin{solution}

\end{solution}

\question Line 355, add a sentence explaining that in SymPy str == repr, because in Python repr is used to get an unambiguous valid Python code repre- sentation, while the return value of str is meant to be human-readable.
\begin{solution}

\end{solution}

\question Lines 359 and 379: what is 2D text representation? It seems that “2D” shouldn’t be here.
\begin{solution}

\end{solution}

\question Line 427: every dictionary is a “dictionary of keys”. This should be a dictionary with coordinate tuples as keys associated with the appropriate values.
\begin{solution}

\end{solution}

\question Section 4 would benefit from an introduction, and lines 440– should be- come a new subsection 4.1. (named “Float” or “Real numbers support” or similar).
\begin{solution}

\end{solution}

\question I suggest a better example for lines 459–460: a computation of (e100 +1)− e100 :
\begin{solution}

\end{solution}
   >>> (exp(100)+1).evalf() - exp(100).evalf()
   0
   >>> ((exp(100)+1) - exp(100)).evalf()
   1.00000000000000
   >>> (exp(100)+1) - exp(100)
   1
or two different ways to compute the 100th Fibonacci number:
   >>> phi = (1+sqrt(5))/2
   >>> psi = (1-sqrt(5))/2
   >>> ((phi**100-psi**100)/sqrt(5)).evalf()-fibonacci(100)
   65536.0000000000
   >>> ((phi**100-psi**100)/sqrt(5) - fibonacci(100)).evalf()
   0.e-104
Obviously, like your own example, these are problematic because a part of the computation is relying on Python’s builtin floats. Please include a comment on whether symbolic computation (i.e., applying evalf() on the whole expression) always avoids these errors or not, possibly with an example when it doesn’t resolve this problem.

\question The footnote from line 477 should be moved as a sentence in its own right to the introduction, with other technical specifications.
\begin{solution}

\end{solution}

\question In line 495, the word “solving” seems more appropriate than “solutions”.
\begin{solution}

\end{solution}

\question In line 504 “is” should be used instead of “are” (because “array” is singular).
\begin{solution}

\end{solution}

\question In line 518, “produces” should become plural.
\begin{solution}

\end{solution}

\question The title of section 5, “Domain specific submodules”, seems inappropriate because the section only covers Physics package (not “submodules”). It should either be expanded with a short introduction listing other domain specific packages, or it should be renamed to “Physics package”.
\begin{solution}

\end{solution}

\question The word “symbolics” in line 528 should be removed (as almost everything in the paper deals with symbolic computation).
\begin{solution}

\end{solution}

\question In line 530, sympy.physics.vector is a module, not a package.
\begin{solution}

\end{solution}

\question It is unclear what “both of these objects” refer to in line 532. My guess is vectors and dyadic objects, but this should be reworded to make it more clear.
\begin{solution}

\end{solution}

\question In lines 543 and 545, “rad” should be removed. Radians are assumed when no other measure (like degrees) is given.
\begin{solution}

\end{solution}

\question In lines 567–568, “performing symbolic quantum mechanics” makes no sense. This should probably be “computations”, “solving problems related to”, etc.
\begin{solution}

\end{solution}

\question The sentence “SymPy expressions are immutable trees of Python objects.” doesn’t belong in the conclusion. This can be moved to the appropriate place when discussing SymPy’s architecture.
\begin{solution}

\end{solution}

\question All “submodules” should be replaced by “modules” (examples: lines 657, 662).
\begin{solution}

\end{solution}

\question The sentences in lines 657–661 should swap places, because “areas of mathematics” are discrete mathematics, concrete mathematics, etc., while simplifying expressions, performing common calculus operations, pretty print- ing expressions, etc. belongs to common operations (“other areas” is also fine, albeit slightly wrong).
\begin{solution}

\end{solution}

\question In line 662 “classical mechanics and quantum mechanics” are listed as the only example of the support for specific domains, as in section 5, which leaves the impression that physics the only one. Either more domains should be listed, or it should be reworded to recognize the fact that there are no others.
\begin{solution}

\end{solution}

\question Lines 670–678 contain explanations what some of the authors’ institu- tions are. It is customary to use acknowledgements to thank people and institutions, while institutions’ details should be provided in the authors’ footnotes in the documents’ head.
\begin{solution}

\end{solution}

\question The citation [5] in line 691 is missing identification data, probably a URL.
\begin{solution}

\end{solution}

\question The citation [21] in line 728 should have “2D” instead of “2d”. More specific comments and correction suggestions for the supplement fol- low.
\begin{solution}

\end{solution}

\question The supplement should have an introduction, explaining what is being covered in it in general and in each section.
\begin{solution}

\end{solution}

\question Since the Guntz algorithm is covered in depth, it would be good to
include how an interested reader can see SymPy’s steps of computation:

\begin{verbatim}
     >>> import os
     >>> os.environ['SYMPY_DEBUG'] = 'True'
     >>> from sympy import *
     >>> x = symbols(’x’)
     >>> limit(sin(x)/x, x, 0)
     DEBUG: parsing of expression [(0, 1, None, None)] with symbol _w
     DEBUG: returned None
     DEBUG: parsing of expression [(_w, 1, None, None)] with symbol _w
     DEBUG: returned ([], [(_w, 1, None, None)], 1, False)
     DEBUG: parsing of expression [(0, 1, None, None)] with symbol _w
     DEBUG: returned None
     DEBUG: parsing of expression [(_w, 1, None, None)] with symbol _w
     DEBUG: returned ([], [(_w, 1, None, None)], 1, False)
     DEBUG: parsing of expression [(_w, 1, None, None)] with symbol _w
     DEBUG: returned ([], [(_w, 1, None, None)], 1, False)
     DEBUG: parsing of expression [(0, 1, None, None)] with symbol _w
     DEBUG: returned None
     DEBUG: parsing of expression [(_w, 1, None, None)] with symbol _w
     DEBUG: returned ([], [(_w, 1, None, None)], 1, False)
     DEBUG: parsing of expression [(0, 1, None, None)] with symbol _w
     DEBUG: returned None
     DEBUG: parsing of expression [(_w, 1, None, None)] with symbol _w
     DEBUG: returned ([], [(_w, 1, None, None)], 1, False)
     DEBUG: parsing of expression [(1, 1, None, None)] with symbol _w
     DEBUG: returned None
     limitinf(x*sin(1/x), x) = 1
     +-mrv_leadterm(_p*sin(1/_p), _p) = (1, 0)
     | +-mrv(_p*sin(1/_p), _p) = ({_p: _Dummy_14}, {}, _Dummy_14*sin(1/_Dummy_14))
     | | +-mrv(_p, _p) = ({_p: _Dummy_14}, {}, _Dummy_14)
     | | +-mrv(sin(1/_p), _p) = ({_p: _Dummy_15}, {}, sin(1/_Dummy_15))
     | |   +-mrv(1/_p, _p) = ({_p: _Dummy_15}, {}, 1/_Dummy_15)
     | |     +-mrv(_p, _p) = ({_p: _Dummy_15}, {}, _Dummy_15)
     | +-rewrite(_Dummy_14*sin(1/_Dummy_14), {exp(_p): _Dummy_14}, {}, _p, _w)
             = (sin(_w)/_w, -_p)
     | | +-sign(_p, _p) = 1
     | | +-limitinf(1, _p) = 1
     | +-calculate_series(sin(_w)/_w, _w) = 1
     |   +-limitinf(_w, _w) = oo
     |   | +-mrv_leadterm(_w, _w) = (1, -1)
     |   | | +-mrv(_w, _w) = ({_w: _Dummy_18}, {}, _Dummy_18)
     |   | | +-rewrite(_Dummy_18, {exp(_w): _Dummy_18}, {}, _w, _w) = (1/_w, -_w)
| |||+-sign(_w,_w)=1
     |   | | | +-limitinf(1, _w) = 1
     |   | | +-calculate_series(1/_w, _w) = 1/_w
     |   | +-sign(-1, _w) = -1
     |   | +-sign(1, _w) = 1
     |   +-limitinf(_w, _w) = oo
     |   +-limitinf(_w, _w) = oo
     |   +-limitinf(_w, _w) = oo
     +-sign(0, _p) = 0
     +-limitinf(1, _p) = 1
\end{verbatim}

Emphasize that the environment variable \texttt{SYMPY\_DEBUG} must be set before importing SymPy for the first time.
\begin{solution}

\end{solution}

\question In line 183, syntax symbols(a:d) is used without being previously defined or explained.
\begin{solution}

\end{solution}

\question Figure 2 should not be a screenshot, but rather proper code with LATEX-
w rendered results and explanations, as was done in the rest of the paper. A
 shortened link to SymPy Gamma for the example’s expression can be included
 for the user to try this for themselves.
\begin{solution}

\end{solution}

\end{questions}

### 2016-08-15 rebuttal.tex

\pagebreak

\section{Comments from Editor}

\begin{questions}
\question I found the \url{http://live.sympy.org/} site very convenient for trying things
out in SymPy, but it is only mentioned in the supplement, which might be
overlooked by readers. I suggest moving that section into the main paper, as
it provides a great way to play along with the examples while reading the
paper.
\begin{solution}
  % TODO
\end{solution}
\question The Basic Usage section omits what I think is an important point:
how does one distinguish between evaluating exp(1) in Python and exp(1) in
SymPy? In other words, how are symbolic constants specified?
\begin{solution}
  % TODO. I'm unclear what the editor is asking here.
\end{solution}
\question The first
thing I tried after seeing Table 2 (simplication functions) did not work:

\begin{verbatim}
>>> trigsimp (exp( Matrix(2, 2, [0, -y, y, 0])) )
\end{verbatim}
fails to recognize cos and
sin. Am I expecting too much?
\begin{solution}
  This indeed does not work. The implementation in \verb|trigsimp| primarily
  thus far has focused on transforming trigonometric functions into other
  trigonometric functions. The ability to simplify complex exponentials is
  something that we would like to work.
  \href{https://github.com/sympy/sympy/issues/11459}{Issue 11459} in our
  public issue tracker tracks this feature.
\end{solution}
\question Section 3.6: are eigenvalues and singular
values included? If not, why not?
\begin{solution}
  % TODO
\end{solution}
\question Like one of the referees, I expected to
see MPFR mentioned in Section 4.1, at least to mention the pros and cons and
say why it isn't used.
\begin{solution}
  % TODO
\end{solution}
\question Section 4.1: please state whether the syntax for
functions in mpmath is identical, or not, to that in Sympy.
\begin{solution}
  % TODO
\end{solution}
\question Unless I have
missed something, a weakness of the paper is that speed is mentioned only
twice, on lines 496 and 666. My assumption is that SymPy is slow compared with
its commercial competitors. Please comment on the speed issue, and not just in
the conclusions.
\begin{solution}
  % TODO
\end{solution}
\question Editors are needed for [20].
\begin{solution}
  % TODO: I do not know how to make bibtex (with apalike) show the editor
  % here.
\end{solution}
\end{questions}

@@

\begin{solution}
There is no simple way to access this in code (the simplest way is
\texttt{sys.\allowbreak{}modules[\allowbreak{}\textquotesingle{}sympy.core.assumptions\textquotesingle{}].\allowbreak{}\_assume\_defined}, which works
around some technical issues with Python's import system, and uses a
private method). The recommended list is in the SymPy documentation, at
\url{http://docs.sympy.org/latest/modules/core.html#module-sympy.core.assumptions}.
\end{solution}
\question Section 2.4
Line 216: This is due in part because the same language, Python, is used both for the internal implementation and the external usage by users.

This reads badly. Perhaps the following might be better?

This is due, in part, to the fact that the same language, Python, is used both for the internal implementation and the external usage by users.

\begin{solution}
We have changed the sentence as suggested.
\end{solution}
\question Line 221:
the phrase `Expression tree' is cited but this is not the first time you've used it. Perhaps cite earlier? Line 129 perhaps?

@@

The measure parameter of the simplify function lets the user specify the Python function used to determine how

\begin{solution}
We have changed the sentence as suggested.
\end{solution}
\question Section 4.1
Should mpmath be cited here?

\begin{solution}
We have added a citation for mpmath.
\end{solution}
\end{questions}

\section{Reviewer 3 (Vedran Sego)}

\begin{questions}

\question Explaining what Python is (lines 70--72) should go before talking about SymPy as a ``CAS written in Python''. Further, the paper assumes a moderate familiarity with Python (for example, Python's console, OOP, and exceptions), and this should be specified. There should be a short note on the used Python console (\texttt{>>>} is the prompt, with the results of computation following immediately in the lines after it). The citation [25] from line 65 should be moved next to ``Python'' in line 70.
\begin{solution}

\end{solution}

\question Line 73 has outdated information. Sage was renamed to SageMath and it no longer aims only at pure mathematics but also at algebra, numerical analysis, etc. The reference [40] should be replaced by a more up-to-date one.
\begin{solution}

\end{solution}

\question The plural ``CASs'' is usually written as ``CASes'' or ``CAS's'', with the latter being somewhat problematic due to it looking like it implies possession.
\begin{solution}

\end{solution}

\question Line 88 mentions ``printers'', but it doesn't state what they are, which is confusing for those readers that are yet to learn the concept in section 3.4.
\begin{solution}

\end{solution}

\question Also in line 88, Jupyter's citation [30] is actually about IPython and should be replaced by a more up-to-date version.
\begin{solution}

\end{solution}

\question The word ``software'' in line 91 is ambiguous; ``library'' or ``package'' would make a better choice.
\begin{solution}

\end{solution}

\question Lines 91--96: ``we discuss/look at/etc'' is the preferred form, instead of ``section discusses/looks at/etc''.
\begin{solution}

\end{solution}

\question The paragraph in lines 103--105 should be moved to the introduction,
and the footnote from line 104 should be added to that paragraph as a
full-blown sentence, expanded by all the relevant technical information
(Python version, OS, \ldots). Given that the end of life for Python 2 is
2020., a comment on whether all the presented examples work in both Python 2
and 3 should be included as well. Further, emphasise that wildcard imports,
``\texttt{import *}'', should almost never be used in programs (see PEP 8, the
item ``wildcard imports''). The same goes for the \texttt{import} mentioned in lines 484--487.
\begin{solution}

\end{solution}

\question Line 119 should be removed, as it is basically a copy of the previous line.
\begin{solution}

\end{solution}

\question In line 121 the word ``stored'' should be replaced by ``used as keys''.
\begin{solution}

\end{solution}

\question What do the authors mean by ``thereby permitting features such as caching'' in line 122? Caching can be done for mutable types as well, just not through hashing.
\begin{solution}

\end{solution}

\question There is no need to repeat ``(CAS)'' in line 124, as it was already given in line 64.
\begin{solution}

\end{solution}

\question In the same line, the word ``represents'' should be replaced by ``stores''.
\begin{solution}

\end{solution}

\question In line 184, ``symbols are'' should be replaced by ``\texttt{t} is'' (the general rule is already given in line 179).
\begin{solution}

\end{solution}

\question The code in line 210 should be made into its own line (like a displaymath formula), for typesetting reasons and better readability.
\begin{solution}

\end{solution}

\question The footnote 4 should be moved from line 221 to line 222, right after ``Basic''.
\begin{solution}

\end{solution}

\question The part ``which defines some basic methods for symbolic expression trees'' should be removed from line 222, as it was already given in line 130.
\begin{solution}

\end{solution}

\question In line 225, the sentence ``Not all SymPy classes are subclasses of \texttt{Expr}.'' sounds confusing as a reader new to SymPy wouldn't expect, for example, symbols to inherit ``Expr''. It would be better to expand this, for example ``Most of the SymPy classes (including \texttt{Symbol}) are subclasses of \texttt{Expr}, but there are exceptions to this rule''.
\begin{solution}

\end{solution}

\question The title ``Features'' in line 276 is ambiguous, as a ``feature''
has no precise meaning in Python (or even software libraries in general). It
should be replaced by ``Packages and modules'' or a similar more precise
wording. The same goes for ``feature'' in most other places in the paper (for
example, the caption of Table 1.\ and line 495). It would be very useful to also include actual names of the packages/modules in Table 1., as well as in any section covering those packages/modules.
\begin{solution}

\end{solution}

\question The sets support listing is unnaturally split in two by the ``This includes\ldots'' sentence which would fit better in parentheses.
\begin{solution}

\end{solution}

\question Line 355, add a sentence explaining that in SymPy \texttt{str == repr}, because in Python \texttt{repr} is used to get an unambiguous valid Python code representation, while the return value of \texttt{str} is meant to be human-readable.
\begin{solution}

\end{solution}

\question Lines 359 and 379: what is 2D text representation? It seems that ``2D'' shouldn't be here.
\begin{solution}

\end{solution}

\question Line 427: every dictionary is a ``dictionary of keys''. This should be a dictionary with coordinate tuples as keys associated with the appropriate values.
\begin{solution}

\end{solution}

\question Section 4 would benefit from an introduction, and lines 440-- should become a new subsection 4.1. (named ``Float'' or ``Real numbers support'' or similar).
\begin{solution}

\end{solution}

\question I suggest a better example for lines 459--460: a computation of $(e^{100} +1)− e^{100}$:

\begin{verbatim}
   >>> (exp(100)+1).evalf() - exp(100).evalf()
   0
   >>> ((exp(100)+1) - exp(100)).evalf()
   1.00000000000000
   >>> (exp(100)+1) - exp(100)
   1
\end{verbatim}

or two different ways to compute the 100th Fibonacci number:

\begin{verbatim}
   >>> phi = (1+sqrt(5))/2
   >>> psi = (1-sqrt(5))/2
   >>> ((phi**100-psi**100)/sqrt(5)).evalf()-fibonacci(100)
   65536.0000000000
   >>> ((phi**100-psi**100)/sqrt(5) - fibonacci(100)).evalf()
   0.e-104
\end{verbatim}

Obviously, like your own example, these are problematic because a part of the computation is relying on Python's builtin floats. Please include a comment on whether symbolic computation (i.e., applying \texttt{evalf()} on the whole expression) always avoids these errors or not, possibly with an example when it doesn't resolve this problem.
\begin{solution}

\end{solution}

\question The footnote from line 477 should be moved as a sentence in its own right to the introduction, with other technical specifications.
\begin{solution}

\end{solution}

\question In line 495, the word ``solving'' seems more appropriate than ``solutions''.
\begin{solution}

\end{solution}

\question In line 504 ``is'' should be used instead of ``are'' (because ``array'' is singular).
\begin{solution}

\end{solution}

\question In line 518, ``produces'' should become plural.
\begin{solution}

\end{solution}

\question The title of section 5, ``Domain specific submodules'', seems inappropriate because the section only covers Physics package (not ``submodules''). It should either be expanded with a short introduction listing other domain specific packages, or it should be renamed to ``Physics package''.
\begin{solution}

\end{solution}

\question The word ``symbolics'' in line 528 should be removed (as almost everything in the paper deals with symbolic computation).
\begin{solution}

\end{solution}

\question In line 530, sympy.physics.vector is a module, not a package.
\begin{solution}

\end{solution}

\question It is unclear what ``both of these objects'' refer to in line 532. My guess is vectors and dyadic objects, but this should be reworded to make it more clear.
\begin{solution}

\end{solution}

\question In lines 543 and 545, ``rad'' should be removed. Radians are assumed when no other measure (like degrees) is given.
\begin{solution}

\end{solution}

\question In lines 567--568, ``performing symbolic quantum mechanics'' makes no sense. This should probably be ``computations'', ``solving problems related to'', etc.
\begin{solution}

\end{solution}

\question The sentence ``SymPy expressions are immutable trees of Python objects.'' doesn't belong in the conclusion. This can be moved to the appropriate place when discussing SymPy's architecture.
\begin{solution}

\end{solution}

\question All ``submodules'' should be replaced by ``modules'' (examples: lines 657, 662).
\begin{solution}

\end{solution}

\question The sentences in lines 657--661 should swap places, because ``areas of mathematics'' are discrete mathematics, concrete mathematics, etc., while simplifying expressions, performing common calculus operations, pretty print- ing expressions, etc.\ belongs to common operations (``other areas'' is also fine, albeit slightly wrong).
\begin{solution}

\end{solution}

\question In line 662 ``classical mechanics and quantum mechanics'' are listed as the only example of the support for specific domains, as in section 5, which leaves the impression that physics the only one. Either more domains should be listed, or it should be reworded to recognize the fact that there are no others.
\begin{solution}

\end{solution}

\question Lines 670--678 contain explanations what some of the authors' institutions are. It is customary to use acknowledgements to thank people and institutions, while institutions' details should be provided in the authors' footnotes in the documents' head.
\begin{solution}

\end{solution}

\question The citation [5] in line 691 is missing identification data, probably a URL\@.
\begin{solution}

\end{solution}

\question The citation [21] in line 728 should have ``2D'' instead of ``2d''. More specific comments and correction suggestions for the supplement fol- low.
\begin{solution}

\end{solution}

\question The supplement should have an introduction, explaining what is being covered in it in general and in each section.
\begin{solution}

\end{solution}

\question Since the Guntz algorithm is covered in depth, it would be good to
include how an interested reader can see SymPy's steps of computation:

\begin{verbatim}
     >>> import os
     >>> os.environ['SYMPY_DEBUG'] = 'True'
     >>> from sympy import *
     >>> x = symbols('x')
     >>> limit(sin(x)/x, x, 0)
     DEBUG: parsing of expression [(0, 1, None, None)] with symbol _w
     DEBUG: returned None
     DEBUG: parsing of expression [(_w, 1, None, None)] with symbol _w
     DEBUG: returned ([], [(_w, 1, None, None)], 1, False)
     DEBUG: parsing of expression [(0, 1, None, None)] with symbol _w
     DEBUG: returned None
     DEBUG: parsing of expression [(_w, 1, None, None)] with symbol _w
     DEBUG: returned ([], [(_w, 1, None, None)], 1, False)
     DEBUG: parsing of expression [(_w, 1, None, None)] with symbol _w
     DEBUG: returned ([], [(_w, 1, None, None)], 1, False)
     DEBUG: parsing of expression [(0, 1, None, None)] with symbol _w
     DEBUG: returned None
     DEBUG: parsing of expression [(_w, 1, None, None)] with symbol _w
     DEBUG: returned ([], [(_w, 1, None, None)], 1, False)
     DEBUG: parsing of expression [(0, 1, None, None)] with symbol _w
     DEBUG: returned None
     DEBUG: parsing of expression [(_w, 1, None, None)] with symbol _w
     DEBUG: returned ([], [(_w, 1, None, None)], 1, False)
     DEBUG: parsing of expression [(1, 1, None, None)] with symbol _w
     DEBUG: returned None
     limitinf(x*sin(1/x), x) = 1
     +-mrv_leadterm(_p*sin(1/_p), _p) = (1, 0)
     | +-mrv(_p*sin(1/_p), _p) = ({_p: _Dummy_14}, {}, _Dummy_14*sin(1/_Dummy_14))
     | | +-mrv(_p, _p) = ({_p: _Dummy_14}, {}, _Dummy_14)
     | | +-mrv(sin(1/_p), _p) = ({_p: _Dummy_15}, {}, sin(1/_Dummy_15))
     | |   +-mrv(1/_p, _p) = ({_p: _Dummy_15}, {}, 1/_Dummy_15)
     | |     +-mrv(_p, _p) = ({_p: _Dummy_15}, {}, _Dummy_15)
     | +-rewrite(_Dummy_14*sin(1/_Dummy_14), {exp(_p): _Dummy_14}, {}, _p, _w)
             = (sin(_w)/_w, -_p)
     | | +-sign(_p, _p) = 1
     | | +-limitinf(1, _p) = 1
     | +-calculate_series(sin(_w)/_w, _w) = 1
     |   +-limitinf(_w, _w) = oo
     |   | +-mrv_leadterm(_w, _w) = (1, -1)
     |   | | +-mrv(_w, _w) = ({_w: _Dummy_18}, {}, _Dummy_18)
     |   | | +-rewrite(_Dummy_18, {exp(_w): _Dummy_18}, {}, _w, _w) = (1/_w, -_w)
     |   | | | +-sign(_w,_w)=1
     |   | | | +-limitinf(1, _w) = 1
     |   | | +-calculate_series(1/_w, _w) = 1/_w
     |   | +-sign(-1, _w) = -1
     |   | +-sign(1, _w) = 1
     |   +-limitinf(_w, _w) = oo
     |   +-limitinf(_w, _w) = oo
     |   +-limitinf(_w, _w) = oo
     +-sign(0, _p) = 0
     +-limitinf(1, _p) = 1
\end{verbatim}

Emphasize that the environment variable \texttt{SYMPY\_DEBUG} must be set before importing SymPy for the first time.
\begin{solution}

\end{solution}

\question In line 183, syntax \texttt{symbols(\textquotesingle{}a:d\textquotesingle{})} is used without being previously defined or explained.
\begin{solution}

\end{solution}

\question Figure 2 should not be a screenshot, but rather proper code with \LaTeX-rendered results and explanations, as was done in the rest of the paper. A
 shortened link to SymPy Gamma for the example's expression can be included
 for the user to try this for themselves.
\begin{solution}

\end{solution}

\end{questions}

### 2016-08-16 rebuttal.tex

\end{solution}

\question Line 73 has outdated information. Sage was renamed to SageMath and it no longer aims only at pure mathematics but also at algebra, numerical analysis, etc. The reference [40] should be replaced by a more up-to-date one.
\begin{solution}

\end{solution}

\question The plural ``CASs'' is usually written as ``CASes'' or ``CAS's'', with the latter being somewhat problematic due to it looking like it implies possession.
\begin{solution}
  % TODO: I recall we looked this up and the spelling consensus agreed with
  % the way we have it. We should double check (perhaps we can email the
  % editor to ask if the journal has a convention).
\end{solution}

@@

\question Also in line 88, Jupyter's citation [30] is actually about IPython and should be replaced by a more up-to-date version.
\begin{solution}
  % See https://github.com/jupyter/jupyter/issues/190
\end{solution}

\question The word ``software'' in line 91 is ambiguous; ``library'' or ``package'' would make a better choice.
\begin{solution}

\end{solution}

\question Lines 91--96: ``we discuss/look at/etc'' is the preferred form, instead of ``section discusses/looks at/etc''.
\begin{solution}
  % TODO: We had decided not to use first person. We should perhaps email the editor
  % to ask about this.
\end{solution}

@@

\question The paragraph in lines 103--105 should be moved to the introduction,
and the footnote from line 104 should be added to that paragraph as a
full-blown sentence, expanded by all the relevant technical information
(Python version, OS, \ldots). Given that the end of life for Python 2 is
2020., a comment on whether all the presented examples work in both Python 2
and 3 should be included as well. Further, emphasise that wildcard imports,
``\texttt{import *}'', should almost never be used in programs (see PEP 8, the
item ``wildcard imports''). The same goes for the \texttt{import} mentioned in lines 484--487.
\begin{solution}

\end{solution}

\question Line 119 should be removed, as it is basically a copy of the previous line.
\begin{solution}
Line 119 is the output of line 118, so now showing it would
be incorrect.

The aim here was simply to show that the input expression remains unevaluated.
We have added a sentence before the example to note this.
This particular example was chosen because it shows the basic syntax for
addition, subtraction, multiplication, division, and exponentiation.
\end{solution}

### 2016-08-16 architecture.tex
Address reviewer 3, comment 9
Move the footnote as suggested by the reviewer
This is a better place for it.

Expressions are created from symbols using Python's mathematical syntax.  For
instance, the following Python code creates the expression $(x^2 - 2x + 3)/y$.

@@

The typical way to create a custom SymPy object is to subclass an existing
SymPy class, usually  \texttt{Basic}, \texttt{Expr}, or
\texttt{Function}. All SymPy classes\footnote{Some
  internal classes, such as those used in the polynomial module, do not follow
  this rule for efficiency reasons.} used for expression trees should be subclasses of the base class
\texttt{Basic}, which defines some basic methods for symbolic expression
trees. \texttt{Expr} is the subclass for mathematical expressions that can be
added and multiplied together. Instances of \texttt{Expr} typically represent
complex numbers, but may also include other ``rings'' like matrix expressions.
Not all SymPy classes are subclasses of \texttt{Expr}. For instance, logic expressions such
as \verb|And(x, y)| are subclasses of \texttt{Basic} but not of \texttt{Expr}.

>>>>

Expressions are created from symbols using Python's mathematical syntax.  For
instance, the following Python code creates the expression $(x^2 - 2x + 3)/y$.
Note that the expression remains unevaluated: it is represented symbolically.

@@

The typical way to create a custom SymPy object is to subclass an existing
SymPy class, usually  \texttt{Basic}, \texttt{Expr}, or
\texttt{Function}. All SymPy classes used for expression trees should be subclasses of the base class
\texttt{Basic}\footnote{Some
  internal classes, such as those used in the polynomial module, do not follow
  this rule for efficiency reasons.}, which defines some basic methods for symbolic expression
trees. \texttt{Expr} is the subclass for mathematical expressions that can be
added and multiplied together. Instances of \texttt{Expr} typically represent
complex numbers, but may also include other ``rings'' like matrix expressions.
Not all SymPy classes are subclasses of \texttt{Expr}. For instance, logic expressions such
as \verb|And(x, y)| are subclasses of \texttt{Basic} but not of \texttt{Expr}.
